---
title: "Lab3_A"
author: "Sanyam Shilpakar"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval= FALSE}
download.file(url = "https://ndownloader.figshare.com/files/2292169",
              destfile = "data_raw/portal_data_joined.csv")
```

```{r}

surveys <- read.csv("data_raw/portal_data_joined.csv")
```

# CONTENT
```{r}
head(surveys)
tail(surveys)
```

```{r, eval= FALSE}
View(surveys)
```

# SUMMARY
```{r}
str(surveys)
summary(surveys)
```
    

# SIZE
```{r}
dim(surveys)
ncol(surveys)
nrow(surveys)
```

# NAMES
```{r}
names(surveys)
rownames(surveys)
```

# CHALLENGE QUESTION 1   
    * the object surveys is of class data.frame.
    * there are 34786 rows and 13 columns in this object.
    * there have been 48 recorded species during this survey.
    
```{r eval = FALSE}
surveys$species
```


# CHALLENGE QUESTION 2
```{r}
surveys_200 <- surveys[1:200,]
nrow(surveys_200)

```

# Indexing and Subsetting
```{r}
surveys_last <- surveys[nrow(surveys_200),]
surveys_last
```

```{r}
tail(surveys_200)
```

```{r}
surveys_middle <- surveys[nrow(surveys_200)/2,]
surveys_middle
```

```{r}
surveys_head <- surveys[-(7: nrow(surveys)), ]
surveys_head
```

# FACTORS
```{r}
sex <- factor(c("male", "female", "female", "male"))
levels(sex)
nlevels(sex)
```

```{r}
sex <- factor(sex, levels = c("male", "female"))
sex
```

# Converting factors
```{r}
as.character(sex)
```

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(year_fct)               # Wrong! And there is no warning...
as.numeric(as.character(year_fct)) # Works...
as.numeric(levels(year_fct))[year_fct]    # The recommended way.
```

# Renaming Factors
```{r}
plot(as.factor(surveys$sex))
```

```{r}
# there are about 1700 individuals for which the sex has not been recorded
sex <- factor(surveys$sex)
head(sex)
```

```{r}
levels(sex)[1] <- "undetermined"
head(sex)
```


# CHALLENGE QUESTION 3
```{r}
levels(sex)[2] <- "female"
levels(sex)[3] <- "male"
head(sex)
```

```{r}
sex <- factor(sex, levels = c("female", "male", "undetermined"))
plot(as.factor(sex))
```

# USING stringAsFactors = FALSE. 
#When set to TRUE, will change characters into factors when you read a CSV file.
```{r}
surveys <- read.csv("data_raw/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
surveys <- read.csv("data_raw/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
## Convert the column "plot_type" into a factor
surveys$plot_type <- factor(surveys$plot_type)
str(surveys)

```

# CHALLENGE QUESTION 4
```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea urchin"),
          feel = c("furry", "furry", "squishy", "spiny"),
          weight = c(45, 8, 1.1, 0.8)
          )
```

  * my prediction is that country, climate, temperature, and northern_hemisphere are factors, whereas, has_kangaroo is a numeric.
  
```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )

str(country_climate)
```

  * it is indeed what I expected.
  * if we had stringsAsFactors = FALSE, all the variables stated would be character vectors instead of factors.
  * removing "" from temperature and northern_hemisphere and replace 1 in has_kangaroo to TRUE would ensure that eacg column had the accurate data type.


