---
title: "lab4"
author: "Sanyam Shilpakar"
date: "9/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```

```{r}
str(report_03_11_2020)
```

```{r}
view(report_03_11_2020)
```

```{r}
library(DT)
datatable(report_03_11_2020)
```

# EXERCISES (PART 1)

#03_13_2020
```{r}
report_03_13_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv"))

view(report_03_13_2020)
str(report_03_13_2020)
```

```{r}
usa_report_03_13_2020 <- report_03_13_2020 %>% 
  filter(`Country/Region` == 'US') %>% 
  group_by(`Country/Region`) %>% 
  summarize(Deaths = sum(Deaths))

usa_report_03_13_2020
```

```{r}
top10_report_03_13_2020 <- report_03_13_2020 %>% 
  group_by(`Country/Region`) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:10)

top10_report_03_13_2020
```

#9/13/2020

```{r}
report_09_13_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) 

view(report_09_13_2020)
str(report_09_13_2020)
```


```{r}
usa_report_09_13_2020 <- report_09_13_2020 %>% 
  filter(`Country_Region` == 'US') %>% 
  group_by(`Country_Region`) %>% 
  summarize(Deaths = sum(Deaths))

usa_report_09_13_2020
```

```{r}
top10_report_09_13_2020 <- report_09_13_2020 %>% 
  group_by(`Country_Region`) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:10)

top10_report_09_13_2020
```


  From the two data analysis, we see a sudden increase in Deaths in US from
  47 deaths in march 13th 2020, to 194,071 deaths in September 13th, 2020. It 
  moved from 7th highest country to 1st in the list of most deaths. Similarly,
  Brazil and India also had extreme increase in death tolls as it categorized 
  as 2nd and 3rd highest death tolls respectively. Overall, the list of highest 
  deaths increased by a lot within 6 months - from approximately 10 - 3200 to approximately 
  23,000 to 194,000.
  

# EXERCISE (PART 2)

```{r}
library(tidyverse)
```

#03/12/2020
```{r}
report_03_12_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-12-2020.csv")) 
```

# 5. 10 countries with deaths in 03/12/2020
```{r}
deaths_top10_report_03_12_2020 <- report_03_12_2020 %>% 
  group_by(`Country/Region`) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:10)

deaths_top10_report_03_12_2020
```


```{r}

deaths_top10_report_03_12_2020 %>% 
  ggplot(mapping = aes(x= `Country/Region`, y =Deaths)) + 
  geom_jitter(alpha = 0.3, aes(color = "red")) +
  geom_point() +
  scale_y_log10() +
  ggtitle("Top 10 Countries with Deaths in 03/12/2020")
``` 

# 5. 10 countries with Confirmed in 03/12/2020
```{r}
confirmed_top10_report_03_12_2020 <- report_03_12_2020 %>% 
  group_by(`Country/Region`) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  arrange(desc(Confirmed)) %>% 
  slice(1:10)

confirmed_top10_report_03_12_2020
```


```{r}

confirmed_top10_report_03_12_2020 %>% 
  ggplot(mapping = aes(x= `Country/Region`, y =Confirmed)) + 
  geom_jitter(alpha = 0.3, aes(color = "red")) +
  geom_point() +
  scale_y_log10() +
  ggtitle("Top 10 Countries with Confirmed in 03/12/2020")
``` 


#09/12/2020
```{r}
report_09_12_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-12-2020.csv")) 
```

# 5. 10 countries with deaths in 09/12/2020
```{r}
deaths_top10_report_09_12_2020 <- report_09_12_2020 %>% 
  group_by(`Country_Region`) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:10)

deaths_top10_report_09_12_2020
```


```{r}

deaths_top10_report_09_12_2020 %>% 
  ggplot(mapping = aes(x= `Country_Region`, y =Deaths)) + 
  geom_jitter(alpha = 0.3, aes(color = "red")) +
  geom_point() +
  scale_y_log10() +
  ggtitle("Top 10 Countries with Deaths in 09/12/2020")
``` 

# 5. 10 countries with Confirmed in 03/12/2020
```{r}
confirmed_top10_report_09_12_2020 <- report_09_12_2020 %>% 
  group_by(`Country_Region`) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  arrange(desc(Confirmed)) %>% 
  slice(1:10)

confirmed_top10_report_09_12_2020
```


```{r}
confirmed_top10_report_09_12_2020 %>% 
  ggplot(mapping = aes(x= `Country_Region`, y =Confirmed)) + 
  geom_jitter(alpha = 0.3, aes(color = "red")) +
  geom_point() +
  scale_y_log10() +
  ggtitle("Top 10 Countries with Confirmed in 09/12/2020")
``` 

# 6. 10 States in 3/12/2020

```{r}
deaths_top10_states_report_03_12_2020 <- report_03_12_2020 %>% 
  group_by(`Province/State`) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:10)

deaths_top10_states_report_03_12_2020
```

```{r}
deaths_top10_states_report_03_12_2020 %>% 
  ggplot(mapping = aes(x= `Province/State`, y =Deaths)) + 
  geom_jitter(alpha = 0.3, aes(color = "red")) +
  geom_bar(stat='identity') +
  ggtitle("Top 10 States with Deaths in 03/12/2020")
``` 

```{r}
confirmed_top10_states_report_03_12_2020 <- report_03_12_2020 %>% 
  group_by(`Province/State`) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  arrange(desc(Confirmed)) %>% 
  slice(1:10)

confirmed_top10_states_report_03_12_2020
```

```{r}
confirmed_top10_states_report_03_12_2020 %>% 
  ggplot(mapping = aes(x= `Province/State`, y =Confirmed)) + 
  geom_jitter(alpha = 0.3, aes(color = "red")) +
  geom_bar(stat='identity') +
  ggtitle("Top 10 States with Confirmed Cases in 03/12/2020")
``` 


# 6. 10 States in 09/12/2020

```{r}
deaths_top10_states_report_09_12_2020 <- report_09_12_2020 %>% 
  group_by(`Province_State`) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:10)

deaths_top10_states_report_09_12_2020
```

```{r}
deaths_top10_states_report_09_12_2020 %>% 
  ggplot(mapping = aes(x= `Province_State`, y =Deaths)) + 
  geom_jitter(alpha = 0.3, aes(color = "red")) +
  geom_bar(stat='identity') +
  ggtitle("Top 10 States with Deaths in 09/12/2020")
``` 

```{r}
confirmed_top10_states_report_09_12_2020 <- report_09_12_2020 %>% 
  group_by(`Province_State`) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  arrange(desc(Confirmed)) %>% 
  slice(1:10)

confirmed_top10_states_report_09_12_2020
```

```{r}
confirmed_top10_states_report_09_12_2020 %>% 
  ggplot(mapping = aes(x= `Province_State`, y =Confirmed)) + 
  geom_jitter(alpha = 0.3, aes(color = "red")) +
  geom_bar(stat='identity') +
  ggtitle("Top 10 States with Confirmed Cases in 09/12/2020")
``` 
