---
title: "lab6"
author: "Sanyam Shilpakar"
date: "9/28/2020"
output: html_document
bibliography: data/references.bib
---

Application written in R [@JHUCSSE] using the Shiny framework [@JHU]

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
```


```{r}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") 
```


```{r}
head(daily_report)
```


```{r}
ggplot(daily_report, aes(x= Long, y= Lat, size= Confirmed/1000))+
  borders("world", color= NA, fill= "grey90") +
  theme_bw()+
  geom_point(shape= 21, color= 'purple', fill= 'purple', alpha= 0.5, size= 1)+
  labs(titles= 'World COVID-19 Confirmed Cases', x= '', y= '', size= "Cases (x1000))")+
  theme(legend.position = "right") +
  coord_fixed(ratio= 1.5)
```


# Exercise 1

```{r}
daily_report_9_26_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") 
```

```{r}
head(daily_report_9_26_2020)
```


```{r}
ggplot(daily_report_9_26_2020, aes(x= Long, y= Lat, size= Deaths/1000))+
  borders("world", color= NA, fill= "grey90") +
  theme_bw()+
  geom_point(shape= 21, color= 'purple', fill= 'purple', alpha= 0.5, size= 1)+
  labs(titles= 'World COVID-19 Deaths: 9/26/2020', x= '', y= '', size= "Counts (x1000))")+
  theme(legend.position = "right") +
  coord_fixed(ratio= 1.1)
```



```{r}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-05-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
```


```{r}
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5, size= 1) +
    labs(title = 'COVID-19 Confirmed Cases in the US', x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```


```{r}
mybreaks <- c(1, 100, 1000, 10000, 10000)
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7, size= 2.5) +
    scale_size_continuous(name="Cases", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "Anisa Dhana's lagout for COVID-19 Confirmed Cases in the US'") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```


# Exercise 2

```{r}
daily_report_9_26_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
```


```{r}
mybreaks <- c(1, 100, 1000, 10000, 10000)
ggplot(daily_report_9_26_2020, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7, size= 2) +
    scale_size_continuous(name="Counts", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Counts",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "Anisa Dhana's lagout for COVID-19 Confirmed Cases in the US: 9/26/2020") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
  
```



```{r}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Province_State = tolower(Province_State))
# load the US map data
us <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
state_join <- left_join(us, daily_report, by = c("region" = "Province_State"))
# plot state map
```

```{r}
head(us)
```


```{r}
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in the US'")
```


```{r}
library(RColorBrewer)
```


```{r}
report_03_27_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))

us <- map_data("state")
counties <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)

state_join <- left_join(counties, report_03_27_2020, by = c("Key"))
```


```{r}
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) +
  
  borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "PuRd"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County") +
  theme_bw() 
```


# Exercise 3

```{r}
report_09_26_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))

head(report_09_26_2020)

us_9_26_2020 <- map_data("state")
counties_9_26_2020 <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)

state_join <- left_join(counties_9_26_2020, report_09_26_2020, by = c("Key"))
```


```{r}
ggplot(data = us_9_26_2020, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) +
  
  borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "GnBu"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County: 9/26/2020") +
  theme_dark() 
```



```{r}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Massachusetts") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
ma_us <- subset(us, region == "massachusetts")
counties <- map_data("county")
ma_county <- subset(counties, region == "massachusetts")
state_join <- left_join(ma_county, daily_report, by = c("subregion" = "Admin2")) 
```


```{r}
ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 

  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colors = brewer.pal(n = 5, name = "BuGn"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in Massachusetts'")
```

```{r}
head(ma_county)
```


```{r}
library(plotly)
ggplotly(
  ggplot(data= ma_county, aes(x= long, y= lat, group= group))+
    coord_fixed(1.3)+
    geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in MA") +
    ggtitle("COVID-19 Cases in MA")+
    labs(x=NULL, y=NULL)+
    theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```


# Exercise 4

```{r}
daily_report_9_26_2020_texas <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Texas") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
texas_cali <- map_data("state")
texas_us <- subset(us, region == "texas")
counties_texas <- map_data("county")
texas_county <- subset(counties_texas, region == "texas")
state_join_texas <- left_join(texas_county, daily_report_9_26_2020_texas, by = c("subregion" = "Admin2")) 
```


```{r}
ggplotly(
  ggplot(data= texas_county, aes(x= long, y= lat, group= group))+
    coord_fixed(1.3)+
    geom_polygon(data = state_join_texas, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in Texas: 9/26/2020") +
    ggtitle("COVID-19 Cases in Texas: 9/26/2020")+
    labs(x=NULL, y=NULL)+
    theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```


# Exercise 5



```{r echo=FALSE}
world_data_3 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-01-2020.csv"))

world_data_3 <- world_data_3 %>% 
  group_by(`Country/Region`) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  arrange(desc(Confirmed)) %>% 
  slice(1:10)

head(world_data_3)
```


_------------------------------------------------------------------------------
Here we will be looking at the dynamics of COVID-19 confirmed cases in the 
United States relative to other countries from March through September 2020. 
_------------------------------------------------------------------------------



```{r}
barplot(log2(Confirmed) ~. , data= world_data_3, main= "Top 10 Confirmed Cases in the World: MARCH", col= "orange", las= 2)
```

_------------------------------------------------------------------------------
As we can see in the graph above, in March, the United states does not lie 
within the top 10 Confirmed case in the world. 
_------------------------------------------------------------------------------



```{r echo = FALSE}
world_data_9 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-01-2020.csv"))

world_data_9 <- world_data_9 %>% 
  group_by(`Country_Region`) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  arrange(desc(Confirmed)) %>% 
  slice(1:10)

head(world_data_9)
```




```{r}
barplot(Confirmed ~. , data= world_data_9, main= "Top 10 Confirmed Cases in the World: SEPTEMBER", col= "cornflowerblue", las=2)
```

_------------------------------------------------------------------------------
However, as we can see in the bar plot above, in September, the United States 
is ranked the country with highest number of confirmed cases in the world. So 
where did it all start?
_------------------------------------------------------------------------------



```{r echo= FALSE}
daily_report_US_3 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Province_State = tolower(Province_State))
# load the US map data
us_report_3 <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
us_state_join_3 <- left_join(us_report_3, daily_report_US_3, by = c("region" = "Province_State"))
# plot state map
```

```{r echo= FALSE}
head(daily_report_US_3)
```


_------------------------------------------------------------------------------
Lets look at the Confirmed cases in United States according to each state. 
_------------------------------------------------------------------------------



```{r echo= FALSE}
ggplotly(
ggplot(data = us_report_3, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = us_state_join_3, aes(fill = Confirmed), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Cavalcanti1", 100, type = "continuous")) +
  labs(title = "COVID-19 Confirmed Cases in the US: 3/26/2020") +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())+
  theme_dark()
)
```



_------------------------------------------------------------------------------
Hover through each state to find out the exact counts. Alternatively, we can 
see from the color gradients that there seems to be more reported cases in the 
east coast relative to the west. Now lets look at which specific states 
reported the highest count of confirmed cases.
_------------------------------------------------------------------------------



```{r echo=FALSE}
daily_report_US_3 <- daily_report_US_3 %>% 
  arrange(desc(Confirmed)) %>% 
  slice(1:10)
```



```{r}
barplot(Confirmed ~. , data= daily_report_US_3, main= "Top 10 States with highest Confirmed Cases: MARCH", col= "green", las= 2)
```


_------------------------------------------------------------------------------
Here we can visibly see the difference in ratio between two highest east coast 
state vs west coast states. New York and New Jersey seems to have way higher 
counts than California and Washington. In fact, new Jersey which is the second 
highest state in East coast has more counts than both Washington and California 
combined. Similarly, New York has more counts than all of the remaining 9 
states combined. 

This comes to show that the pandemic in the United States may have started from
the east coast. Now lets see the dynamics of the pandemic in the US over time. 
_------------------------------------------------------------------------------




```{r echo= FALSE}
daily_report_US_9 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Province_State = tolower(Province_State))
# load the US map data
us_report_9 <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
us_state_join_9 <- left_join(us_report_9, daily_report_US_9, by = c("region" = "Province_State"))
# plot state map
```



```{r echo= FALSE}
ggplotly(
ggplot(data = us_report_9, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = us_state_join_9, aes(fill = Confirmed), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Cavalcanti1", 100, type = "continuous")) +
  labs(title = "COVID-19 Confirmed Cases in the US: 9/26/2020") +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())+
  theme_dark()
)
```


_------------------------------------------------------------------------------
The graph above shows us that over time the confirmed cases in US increased at 
a very high rate. Compared to the graph from march, which has majority of 
yellow gradients, now we see that majority have green. Some states like Texas 
and California seems to have one of th highest counts. Lets look at it in more
detail.
_------------------------------------------------------------------------------




```{r echo=FALSE}
daily_report_US_9 <- daily_report_US_9 %>% 
  arrange(desc(Confirmed)) %>% 
  slice(1:10)
```

```{r echo=FALSE}
barplot(Confirmed ~. , data= daily_report_US_9, main= "Top 10 States with highest Confirmed Cases: SEPTEMBER", col= "pink", las= 2)
```

_------------------------------------------------------------------------------
Here we can see the difference between states more precisely. The top 2 states
with the highest counts are California and Texas which are west coast. However,
Florida has experienced a sudden increase in counts (college partying?).


Hence, we can finally conclude the dynamics of the 2020 COVID-19 outbreak in 
the United States. The outbreak mainly started from the east coast earlier in
the year and over time spreaded to the west coast and downwards to florida.
There are various factors behind such a pattern that we see. One of the main 
being the fact that these states were initially low in counts. How does that
matter? States which were considered "safe" initially tend to loosen up in 
terms of social distancing and other safety measures, therefore, resulting
in such a shift.
_------------------------------------------------------------------------------



# REFERENCE


