---
title: "lab9"
author: "Sanyam Shilpakar"
date: "10/26/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 14.3 Matching patterns with regular expressions

```{r}
library(stringr)
library(tidyverse)
```

# 14.3.1

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```


```{r}
x <- c("apple", "banana", "pear")
str_view(x, ".a.")
```


```{r}
dot <- "\\."
writeLines(dot)
```

## looking for an explicit .
```{r}
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```


```{r}
x <- "a\\b"
writeLines(x)
```


## for a literal match to \
```{r}
str_view(x, "\\\\")
```

# 14.3.1.1 Exercise 

"\" does not match \ because \ is used as an escape for a new character in regular expressions. 
"\\" will results in '\' which will then act as an escape for the new character. 
In "\\\", the first two backlashes will result in '\' in a regular expression and the remaining third backlash will act as an escape for the new character.


# 14.3.2 Anchors

"^" to match start of string
"$" to match end of a string

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```


```{r}
str_view(x, "a$")
```


```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```


```{r}
str_view(x, "apple$")
```

# 14.3.2.1 Exercise

```{r}
y <- c("Sanyam$^$, Shil$^$pakar")
str_view(y, "\\$\\^\\$")
```


# 14.3.3 Character classes and alternatives 

```{r}
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```


```{r}
str_view(x, ".[*]c")
```


```{r}
str_view(x, "a[ ]")
```


```{r}
x <- c("color", "colour")
str_view(x, "colo(|u)r")
```

# 14.3.3.1 Exercise

```{r}
z <- c("apple", "sanyam", "cry", "owned", "reed", "coding", "compromise")
```

```{r}
# Start with a vowel.
str_view(z, "^[aeiou]")
```


```{r}
# That only contain consonants. (Hint: thinking about matching “not”-vowels.)
str_view(z, "[aeiou]", match = FALSE)
```


```{r}
# End with ed, but not with eed.
str_view(z, "[^e]ed$")
```


```{r}
# End with ing or ise.
z <- c("reed", "coding", "compromise", "yaisg")
str_view(z, "i(ng|se)$")
```

# 14.3.4 Repetition

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```


```{r}
str_view(x, "CC+")
```

```{r}
str_view(x, 'C[LX]+')
```


```{r}
str_view(x, "C{2}")
```


```{r}
str_view(x, "C{2,}")
```


```{r}
str_view(x, 'C{2,3}?')
```


```{r}
str_view(x, 'C[LX]+?')
```


# 14.3.4.1 Exercise

## Describe the equivalents of ?, +, * in {m,n} form.

? is equivalent to {0, 1} which means match at most 1

+ is equivalent to {1,} which means match 1 or more than 1

* is equivalent to {0,} which means match 0 or more



# 14.3.5 Grouping and backreferences

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

# 14.3.5.1 Exercises

1.  `(.)\1\1` : same character three times in a row.

2.  `"(.)(.)\\2\\1"`: A pair of characters followed by the same pair of characters in reversed order.

3.  `(..)\1`: Any two characters repeated.

4.  `"(.).\\1.\\1"`: A character followed by any character, the original character, any other character, the original character again.

5.  `"(.)(.)(.).*\\3\\2\\1"`: Three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order.


# 14.4 Tools

# 14.4.1 Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```


```{r}
# words that start with t
sum(str_detect(words, "^t"))
```


```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```


```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
```

```{r}
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```


```{r}
words[str_detect(words, "x$")]
```


```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```


```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
sum(str_count(x, "a"))
```


```{r}
mean(str_count(words, "[aeiou]"))
```


```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```


```{r}
str_count("abababa", "aba")
```


```{r}
str_view_all("abababa", "aba")
```


# 14.4.1.1 Exercise

```{r}
words[str_detect(words, "^x|x$")]
```


```{r}
# Find all words that start with a vowel and end with a consonant.
start <- str_detect(words, "^[aeiou]")
end <- str_detect(words,"[^aeiou]$")
words[start & end] 
```

```{r}
# Are there any words that contain at least one of each different vowel?

str_view(words, "aeiou")
```


# 14.4.2 Extract matches

```{r}
length(sentences)
head(sentences)
```


```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```


```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```


```{r}
has_colour
```


```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```


```{r}
str_extract(more, colour_match)
```


```{r}
str_extract_all(more, colour_match)
```


```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```

```{r}
str_extract_all(x, "[a-z]", simplify = TRUE)
```


# 14.4.2.1 Exercises

```{r}
# original color match pattern
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
```


```{r}
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2
```


```{r}
more2 <- sentences[str_count(sentences, colour_match) > 1]
```


```{r}
str_view_all(more2, colour_match2, match = TRUE)
```

# 14.4.3 Grouped matches

```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```


```{r}
has_noun %>% 
  str_match(noun)
```


## tibble lets us name the matches. For ex: article, noun, etc.

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```


# 14.4.3.1 Exercises

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("number", "word"), "(one|two|three|four|five|six|seven|eight|nine|ten) ([^ ]+)",
    remove= FALSE
  )
```

# 14.4.4 Replacing matches

```{r}
# only first vowel of the word
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```


```{r}
# all vowels in a word
str_replace_all(x, "[aeiou]", "-")
```


```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```


```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```


# 14.4.4.1 Exercises

```{r}
str_replace_all("hey/there/how/you/doing", "/", "\\\\")
```


# 14.4.5 Splitting

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```


```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```


```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```


```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```


```{r}
str_split(x, " ")[[1]]
```


```{r}
str_split(x, boundary("word"))[[1]]
```


# 14.4.5.1 Exercise 

```{r}
s <- "apples, pears, and bananas"
str_split(s, boundary("word"))[[1]]
```


